---
title: "MineSDG-introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MineSDG-introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(MineSDG)
```

# Introduction

MineSDG provides tools to explore and retrieve official Sustainable Development Goal (SDG) indicator data from the United Nations SDG API.

The package includes:

 - Metadata exploration
 - Indicator validation
 - Country-level data retrieval
 - Optional data persistence
 - Session-level metadata caching

# Listing SDG Indicators

To view all indicators:

```{r eval=FALSE}
head(list_sdg_indicators())
```

To list indicators under a specific goal:

```{r eval=FALSE}
list_sdg_indicators(goal = 15)
```

# Validating Indicators

```{r eval=FALSE}
validate_sdg_indicator("15.3.1")
```

Goal consistency validation:

```{r eval=FALSE}
validate_sdg_indicator("15.3.1", goal = 15)
```

# Fetching Data for a Single Indicator

```{r eval=FALSE}
dt <- fetch_sdg_country_data(
  indicator = "15.3.1",
  country = "IND",
  year_range = c(2015, 2020)
)

head(dt)
```

# Fetching All Indicators Under a Goal

```{r eval=FALSE}
dt_goal <- fetch_sdg_country_data(
  goal = 15,
  country = "IND",
  year_range = c(2015, 2020)
)

head(dt_goal)
```

# Saving Data

```{r eval=FALSE}
fetch_sdg_country_data(
  indicator = "15.3.1",
  country = "IND",
  year_range = c(2015, 2020),
  save = TRUE
)
```

Files are saved under:

`./data/sdg_downloads/`

