---
title: "MineSDG - SDG Analytics & Mining-Sector Intelligence"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MineSDG - SDG Analytics & Mining-Sector Intelligence}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(MineSDG)
```

------------------------------------------------------------------------

# 1. Introduction

MineSDG provides tools to explore and retrieve official Sustainable Development Goal (SDG) indicator data from the United Nations SDG API.

The package includes:

 - Metadata exploration
 - Indicator validation
 - Country-level data retrieval
 - Optional data persistence
 - Session-level metadata caching

------------------------------------------------------------------------

# 2. SDG Analytics Layer

## Listing SDG Indicators

To view all indicators:

```{r eval=FALSE}
head(list_sdg_indicators())
```

To list indicators under a specific goal:

```{r eval=FALSE}
list_sdg_indicators(goal = 15)
```

------------------------------------------------------------------------

## Validating Indicators

```{r eval=FALSE}
validate_sdg_indicator("15.3.1")
```

Goal consistency validation:

```{r eval=FALSE}
validate_sdg_indicator("15.3.1", goal = 15)
```

------------------------------------------------------------------------

## Fetching Data for a Single Indicator

```{r eval=FALSE}
dt <- fetch_sdg_country_data(
  indicator = "15.3.1",
  country = "IND",
  year_range = c(2015, 2020)
)

head(dt)
```

------------------------------------------------------------------------

## Fetching All Indicators Under a Goal

```{r eval=FALSE}
dt_goal <- fetch_sdg_country_data(
  goal = 15,
  country = "IND",
  year_range = c(2015, 2020)
)

head(dt_goal)
```

------------------------------------------------------------------------

## Saving Data

```{r eval=FALSE}
fetch_sdg_country_data(
  indicator = "15.3.1",
  country = "IND",
  year_range = c(2015, 2020),
  save = TRUE
)
```

Files are saved under:

```{r eval=FALSE}
./data/sdg_downloads/
```

------------------------------------------------------------------------

## Advanced Analytics

MineSDG 0.2.0 expands the package beyond data retrieval into a complete SDG analytics engine.

The following analytical layers are now available:

 -  Trend analysis
 -  Stability and volatility diagnostics
 -  Benchmark comparison
 -  Convergence testing
 -  Executive-ready narrative generation
 -  Visualization utilities

------------------------------------------------------------------------

## Trend Analysis

```{r eval=FALSE}
trend <- analyze_sdg_trend(dt)
trend
```

This function computes:

 -  Absolute change
 -  Percent change
 -  CAGR (Compound Annual Growth Rate)
 -  Linear trend slope
 -  Trend direction (Increasing / Decreasing / Stable)

------------------------------------------------------------------------

## Stability & Volatility Diagnostics

```{r eval=FALSE}
stability <- compute_sdg_stability(dt)
stability
```

Includes:

 -  Standard deviation
 -  Coefficient of variation
 -  Volatility index
 -  Stability classification

These metrics are especially useful for ESG risk evaluation and operational performance diagnostics.

------------------------------------------------------------------------

## Benchmarking Performance

```{r eval=FALSE}
benchmark <- benchmark_sdg_performance(dt)
benchmark
```

Outputs:

 -  Deviation from benchmark
 -  Percent gap
 -  Z-score
 -  Ranking
 -  Performance category (Strong / Moderate / Weak)

------------------------------------------------------------------------

## Convergence Analysis

```{r eval=FALSE}
convergence <- analyze_sdg_convergence(dt)
convergence
```


This tests beta-convergence across countries, helping assess whether
lagging countries are catching up in SDG performance.

------------------------------------------------------------------------

## Executive Summary Generation

```{r eval=FALSE}
summary_text <- generate_sdg_executive_summary(trend, benchmark)
cat(summary_text)
```

Produces narrative, board-ready interpretation suitable for:

 -  ESG reports
 -  Policy briefs
 -  Academic summaries
 -  Strategic reviews

------------------------------------------------------------------------

## Visualization Layer

MineSDG includes publication-ready visualization tools:

```{r eval=FALSE}
plot_sdg_trend(dt)
plot_sdg_benchmark(benchmark)
plot_sdg_volatility(stability)
plot_sdg_convergence(dt)
```

All plots use ggplot2 (optional dependency) and follow minimal, publication-ready styling.

------------------------------------------------------------------------

# 3. Mining-Sector Interpretation Layer

MineSDG now includes a sector-specific interpretation layer tailored for mining sustainability analytics.

This enables translation of SDG indicators into mining-relevant sustainability domains, relevance scoring, and strategic narrative interpretation.

------------------------------------------------------------------------

## Mapping SDG to Mining Domains

```{r eval=FALSE}
map_sdg_to_mining_domain(indicator = "15.3.1")
```

Example output:

 - *Goal:* 15
 - *Domain:* Biodiversity & Land
 - *Relevance Score:* 5
 - *Narrative:* Land degradation, rehabilitation, and biodiversity restoration are core mining sustainability metrics.

------------------------------------------------------------------------

## Domain Categories

The following mining sustainability domains are currently defined:

- Climate & Energy
- Water & Resource Efficiency
- Biodiversity & Land
- Community & Social Impact
- Governance & Institutions
- Economic Development
- Health & Safety
- Education & Workforce

This layer enables sector-aware ESG analytics beyond generic SDG evaluation.

------------------------------------------------------------------------

## Mining Risk Engine

```{r eval=FALSE}
dt <- fetch_sdg_country_data(indicator = "15.3.1", country = "IND")
generate_mining_risk_profile(dt, indicator = "15.3.1")
```

------------------------------------------------------------------------

## Mining Risk Profiling

- Indicator-level risk scoring
- Trend + volatility + benchmark integration
- Domain relevance scoring

Example:

```{r eval=FALSE}
risk_profile <- generate_mining_risk_profile(
  data = dt,
  indicator = "15.3.1"
)

risk_profile
```

------------------------------------------------------------------------

## Composite Mining ESG Index

```{r eval=FALSE}
index <- generate_mining_esg_index(
  data = dt_multi,
  indicators = c("6.4.1", "13.2.2", "15.3.1"),
  weighting_method = "domain_weighted"
)

index
```

------------------------------------------------------------------------

## Plot Mining ESG Index

```{r eval=FALSE}
plot_mining_esg_index(index)
```

------------------------------------------------------------------------
## Summary

MineSDG now supports a full analytical workflow:

Data Retrieval → Trend Diagnostics → Stability Analysis → Benchmarking →
Convergence Testing → Executive Reporting → Visualization

This positions MineSDG as a research-grade SDG analytics framework
suitable for:

 -  ESG reporting
 -  Policy evaluation
 -  Academic research
 -  Mining-sector sustainability assessment
